### 2016-08-31 version0.0/intro0.01.tex
Kinda adding why SOM and relative differences.

\section{Prelude}
\label{sec: overview}
%what doIhave inside galaxies:
% stuffIdon't care about because they don't have a big contribution on the SED 1) planets; 2) black holes 3) dark matter
Inside every galaxy, there are billions of stars, planets, and black holes, as well as dust and gas.
They all affect each others' formation, evolution, and extinction.
There is also dark matter inside every galaxy, which, to the best of our knowledge, only has gravitational effects on the other components.
Nearly all the information we can obtain from galaxies other than our own comes from their electromagnetic radiation.
Since each observable phenomenon inside galaxies emits most of its energy in specific wavelengths, 
a plot of brightness/flux density of galaxies as a function of wavelength, called a spectral energy distribution (SED), is a useful tool for understanding galaxies. 
Young and evolved stars, dust, and gas in the space between the stars are the most important contributors to the spectral energy distributions of galaxies.
Planets and black holes without an accretion disc, on the other hand, have a much lower intensity compared to the other aforementioned components, and therefore contribute less to the SED.
Therefore, by properly modeling of observed features in SEDs we can extract information about the evolution of stars, dust, and gas inside galaxies.

@@

%ISM 1)Gas 2)dust
%There could be a galaxy without stars (i.e. dark galaxies~\citep[][and references therein]{Cantalupo12}), but there will never be a galaxy without gas. %: this is a very strong statement. Some galaxies, like dwarf ellipticals, are pretty gas-free (at least now). So qualify this a bit. done
Among observable objects in galaxies, gas has a crucial role in formation and evolution of galaxies.
Gas and dust in galaxies can be found in the space between stars.
This space is called the interstellar medium (ISM), and is full of low density gaseous clouds with neutral or ionized atoms and/or molecules, and microscopic dust grains.
The gas and dust in the ISM are heated by the interstellar radiation field and cooled through a variety of line and continuum processes which usually depend on the local physical conditions. 
The effect of the ISM on stellar formation is undeniable; it provides the raw material for the formation of stars~\citep[e.g.][]{Kennicutt08,Bigiel08}.
Studying the interstellar medium is not only essential to understanding the formation of stars, but is also  necessary for interpretation of stellar evolution due to fact that stars release their material into ISM when they reach the end of their evolutionary track.

In a galaxy like our own, star formation begins with the condensation of giant molecular clouds (GMCs, $\sim 10^7$ M$_{\odot}$) in the ISM due to gravitational instabilities. 
The denser regions within GMCs collapse under their own gravity (self-gravity) and create clumps.
Some of these clumps, which have a similar mass distribution to the stellar initial mass function (IMF), develop into self-gravitating cores.
Then these clumps continue to become denser and turn into protostars. 
Since the protostars have a higher mass than their surroundings, more gas from GMCs accretes onto protostars.
This procedure continues until hydrogen burning in centre of protostars starts (see~\cite{McKee07} for more detail).

Metallicity, which is defined as the ratio of abundance of metals (any element heavier than hydrogen and helium) to the abundance of hydrogen, affects formation and evolution of stars indisputably.
\cite{Walch11} showed that in the case of large scale turbulence, the effect of metallicity on star formation is negligible, but if turbulence is decaying,  metallicity has a strong impact on stabilizing self-gravitating cores.
Metals also play an important role in heating, cooling, and ionizing processes in the ISM.
Since cooling is one of the main reasons for condensation of GMCs~\citep[e.g.][]{Maio07}, in low metallicity regions star formation may be suppressed. 
The amount of metals in stars affects their evolutionary tracks~\citep[e.g.][and references therein]{Maeder02}.
For instance, low metallicity stars tend to have lower mass loss via stellar winds than those with higher metallicity.
Observations show that massive low metallicity stars rotate faster.
Being more massive at the end of their evolutionary tracks gives massive low metallicity stars have a lower chance to explode and become a supernova.

@@

 %how do you model it? %SED model CIGALE %Population synthesis ----> SFR, stellar mass 
The stellar population synthesis method, which sums stellar spectra to produce a composite spectrum, has been widely utilized since the 1970s~\citep[e.g.][]{Tinsley72,Searle73} to model SEDs of galaxies.
A thorough SED model must include stellar population models~\citep[e.g.][]{Bruzual93,Bruzual03,Maraston05}, stellar emission and dust attenuation~\citep[e.g.][]{Calzetti00,Dopita05}, dust grain emission such as from polycyclic aromatic hydrocarbons~\citep[PAHs; e.g.][and references therein]{Tielens08}, and IR emission from gas and dust~\citep[e.g.][]{Chary01,Dale02,Lagache03,Lagache04,Smith07a,Draine07}.
Many groups have modelled SEDs for different types of galaxies and created SED templates (for more information see the review of \cite{Walcher11} on fitting of SEDs and references therein).
Since the physical parameters of templates generated by the SED models are known, we can find properties of observed galaxies by fitting their SEDs (either by minimizing $\chi^2$ or by Bayesian analysis), using the templates.
One example of an SED-fitting code is {\em CIGALE}: Code Investigating GALaxy Emission~\citep{Noll09}.
It combines SED models and uses Bayesian analysis to find the best-fit SED for observed data in the rest-frame UV-to-IR.
Some of the physical properties of galaxies must be assumed as model input find the best SED, and others are derived from SED-fitting results~\citep[see][for more detail]{Walcher08}.
For example, the best-fit SED has information regarding star formation history and stellar mass to light ratio.
By integrating instantaneous star formation over a specific time scale, SFR can be derived.
Knowing the stellar mass to light ratio, one can use a galaxy's observed luminosity to calculate its stellar mass.

@@

\section{Components of the Interstellar Medium} 
\label{sec: ism_intro}
%how do I observe it
%what the observation tell us
% how do I measure the dust and gas
The interstellar medium in galaxies is a complex system with a wide range of structures, but these structures can be divided into two main components: gas and dust.
The gas in the ISM of galaxies is mostly made of hydrogen and helium, with 1 per cent heavier elements (metals), and can be in the form of clouds or diffused regions.
Gas clouds include molecular clouds, cold neutral atomic hydrogen (\hi) or hot \hi~clouds, and diffuse gas can be either \hii~(ionized hydrogen) regions or a hot intercloud medium.
Dust is mostly comprised of silicates, and graphite grains can be seen in dark clouds or galactic cirrus.
Emission lines from atoms or ions (i.e \hii~regions) in a hot gas, lines from molecules in cold clouds, and 21~cm emission of \hi~are the main observational signatures of the gas in the ISM of galaxies.
Dust can be observed directly through far-infrared and sub-millimeter wavelength emission or indirectly through its extinction and attenuation effects (see Section~\ref{sec: extinction}).

@@

The molecular clouds are surrounded by clouds of atomic gas~\citep{Kennicutt12}, which are easily traceable by 21-cm emission.
Hydrogen atoms in the ISM (excluding the regions close to hot stars, see below) are in their ground state, which is the state of lowest energy.
Due to a coupling between the total angular momentum of the electron and the angular momentum (spin) of the nucleus, the ground state of hydrogen atom splits into two states that are separated in energy by an amount of 9.4E-25~J.
This splitting is referred to as the hyperfine structure. 
The electron can be in either of the two aforementioned states. 
These states are very stable and the transition between them is forbidden, with a transition rate of $2.5\times10^{-15}\mathrm{s}^{-1}$.
This means that one would expect this transition to naturally happen once every ten million years.
%The electron and proton spins in the ground state can be parallel (i.e. both spin up) or anti-parallel (i.e. the proton's spin is up and the electron's spin is down or vice versa). 
%The energy of the anti-parallel mode is slightly less than the parallel mode.
%Since atoms always tend to be in the state with lowest energy possible, electrons in the parallel mode have a tendency to flip to the anti-parallel mode. 
%However, the energy difference between these two states is very small and it takes millions of years before the transition happens.
However, the large amounts of hydrogen atoms in these atomic clouds compensate for the rarity of the transition and at any given time, there are enough hydrogen atoms to emit the 21-cm radiation. 
Observation of the \hi~atoms is specifically necessary for understanding physical properties and dynamics of the ISM, which leads to understanding the star formation process~\citep{Walter08}.

>>>>

\section{Prelude}
\label{sec: overview}
%what doIhave inside galaxies:
% stuffIdon't care about because they don't have a big contribution on the SED 1) planets; 2) black holes 3) dark matter
Inside every galaxy, there are billions of stars, planets, and black holes, as well as dust and gas.
They all affect each others' formation, evolution, and extinction.
There is also dark matter inside every galaxy, which, to the best of our knowledge, only has gravitational effects on the other components.
Nearly all the information we can obtain from galaxies other than our own comes from their electromagnetic radiation.
Since each observable phenomenon inside galaxies emits most of its energy in specific wavelengths, 
a plot of brightness/flux density of galaxies as a function of wavelength, called a spectral energy distribution (SED), is a useful tool for understanding galaxies. 
Young and evolved stars, dust, and gas in the space between the stars are the most important contributors to the spectral energy distributions of galaxies.
Planets and black holes without an accretion disc, on the other hand, have much lower luminosities compared to the other aforementioned components, and therefore contribute less to the SED.
Therefore, by properly modeling of observed features in SEDs we can extract information about the evolution of stars, dust, and gas inside galaxies.

@@

%ISM 1)Gas 2)dust
%There could be a galaxy without stars (i.e. dark galaxies~\citep[][and references therein]{Cantalupo12}), but there will never be a galaxy without gas. %: this is a very strong statement. Some galaxies, like dwarf ellipticals, are pretty gas-free (at least now). So qualify this a bit. done
Among observable objects in galaxies, gas has a crucial role in formation and evolution of galaxies.
Gas and dust in galaxies can be found in the space between stars.
This space is called the interstellar medium (ISM), and it is full of low density gaseous clouds with neutral or ionized atoms and/or molecules, and microscopic dust grains.
The gas and dust in the ISM are heated by the interstellar radiation field and cooled through a variety of line and continuum processes which usually depend on the local physical conditions. 
The effect of the ISM on stellar formation is undeniable; it provides the raw material for the formation of stars~\citep[e.g.][]{Kennicutt08,Bigiel08}.
Studying the interstellar medium is not only essential to understanding the formation of stars, but is also  necessary for interpretation of stellar evolution due to fact that stars release their material into ISM when they reach the end of their evolutionary track.

In a galaxy like our own, star formation begins with the condensation of giant molecular clouds (GMCs, $M \sim 10^7$ M$_{\odot}$) in the ISM due to gravitational instabilities. 
The denser regions within GMCs collapse under their own gravity (self-gravity) and create clumps.
Some of these clumps, which have a similar mass distribution to the stellar initial mass function (IMF), develop into self-gravitating cores.
Then these clumps continue to become denser and turn into protostars. 
Since the protostars have a higher mass than their surroundings, more gas from GMCs accretes onto protostars.
This procedure continues until hydrogen burning in the centres of protostars starts (see~\cite{McKee07} for more detail).

Metallicity, which is defined as the ratio of abundance of metals (any element heavier than hydrogen and helium) to the abundance of hydrogen, affects formation and evolution of stars indisputably.
\cite{Walch11} showed that in the case of large scale turbulence, the effect of metallicity on star formation is negligible, but if turbulence is decaying,  metallicity has a strong impact on stabilizing self-gravitating cores.
Metals also play an important role in heating, cooling, and ionizing processes in the ISM.
Since cooling is one of the main reasons for condensation of GMCs~\citep[e.g.][]{Maio07}, in low metallicity regions star formation may be suppressed. 
The amount of metals in stars affects their evolutionary tracks~\citep[e.g.][and references therein]{Maeder02}.
For instance, low metallicity stars tend to have lower mass loss via stellar winds than those with higher metallicity.
Observations show that massive low metallicity stars rotate faster.
Being more massive at the end of their evolutionary tracks gives massive low metallicity stars a lower chance to explode and become a supernova.

@@

 %how do you model it? %SED model CIGALE %Population synthesis ----> SFR, stellar mass 
The stellar population synthesis method, which sums stellar spectra to produce a composite spectrum, has been widely utilized since the 1970s~\citep[e.g.][]{Tinsley72,Searle73} to model SEDs of galaxies.
A thorough SED model must include stellar population models~\citep[e.g.][]{Bruzual93,Bruzual03,Maraston05}, stellar emission and dust attenuation~\citep[e.g.][]{Calzetti00,Dopita05}, dust grain emission such as from polycyclic aromatic hydrocarbons~\citep[PAHs; e.g.][and references therein]{Tielens08}, and IR emission from gas and dust~\citep[e.g.][]{Chary01,Dale02,Lagache03,Lagache04,Smith07a,Draine07}.
Many groups have modelled SEDs for different types of galaxies and created SED templates (for more information see the review of \cite{Walcher11} on fitting of SEDs and references therein).
Since the physical parameters of templates generated by the SED models are known, we can find properties of observed galaxies by fitting their SEDs (either by minimizing $\chi^2$ or by Bayesian analyses), using the templates.
One example of an SED-fitting code is {\em CIGALE}: Code Investigating GALaxy Emission~\citep{Noll09}.
It combines SED models and uses Bayesian analysis to find the best-fit SED for observed data in the rest-frame UV-to-IR.
Some of the physical properties of galaxies must be assumed as model input find the best SED, and others are derived from SED-fitting results~\citep[see][for more detail]{Walcher08}.
For example, the best-fit SED has information regarding star formation history and stellar mass to light ratio.
By integrating instantaneous star formation over a specific time scale, SFR can be derived.
Knowing the stellar mass to light ratio, one can use a galaxy's observed luminosity to calculate its stellar mass.

@@

\section{Components of the Interstellar Medium} 
\label{sec: ism_intro}
%how do I observe it
%what the observation tell us
% how do I measure the dust and gas
The interstellar medium in galaxies is a complex system with a wide range of structures, but these structures can be divided into two main components: gas and dust.
The gas in the ISM of galaxies is mostly made of hydrogen and helium, with 1 per cent heavier elements (metals), and can be in the form of clouds or diffused regions.
Gas clouds include molecular clouds, cold neutral atomic hydrogen (\hi) or hot \hi~clouds, and diffuse gas can be either \hii~(ionized hydrogen) regions or a hot intercloud medium.
Dust is mostly comprised of silicates, and graphite grains can be seen in dark clouds or galactic cirrus.
Emission lines from atoms or ions (i.e.\ \hii~regions) in a hot gas, lines from molecules in cold clouds, and 21~cm emission of \hi~are the main observational signatures of the gas in the ISM of galaxies.
Dust can be observed directly through far-infrared and sub-millimeter wavelength emission or indirectly through its extinction and attenuation effects (see Section~\ref{sec: extinction}).

@@

The molecular clouds are surrounded by clouds of atomic gas~\citep{Kennicutt12}, which are easily traceable by 21-cm emission.
Hydrogen atoms in the ISM (excluding the regions close to hot stars, see below) are in their ground state, which is the state of lowest energy.
Due to a coupling between the total angular momentum of the electron and the angular momentum (spin) of the nucleus, the ground state of hydrogen atom splits into two states that are separated in energy by an amount of $9.4\times 10^{-25}$~J.
This splitting is referred to as the hyperfine structure. 
The electron can be in either of the two aforementioned states. 
These states are very stable and the transition between them is forbidden, with a transition rate of $2.5\times10^{-15}$~s$^{-1}$.
This means that one would expect this transition to naturally happen once every ten million years.
%The electron and proton spins in the ground state can be parallel (i.e. both spin up) or anti-parallel (i.e. the proton's spin is up and the electron's spin is down or vice versa). 
%The energy of the anti-parallel mode is slightly less than the parallel mode.
%Since atoms always tend to be in the state with lowest energy possible, electrons in the parallel mode have a tendency to flip to the anti-parallel mode. 
%However, the energy difference between these two states is very small and it takes millions of years before the transition happens.
However, the large amounts of hydrogen atoms in these atomic clouds compensate for the rarity of the transition and at any given time, there are enough hydrogen atoms to emit the 21-cm radiation. 
Observation of the \hi~atoms is specifically necessary for understanding physical properties and dynamics of the ISM, which leads to understanding the star formation process~\citep{Walter08}.

### 2016-08-31 version0.0/paper3.tex
Kinda adding why SOM and relative differences.

Some of the most tested methods of data mining in astronomy are artificial neural networks (ANNs)~\citep[e.g.][and references therein]{Hossein14,Hossein16}.
ANNs are inspired to work in the same way that the neurons work in a human brain.
These are networks of interconnected neurons (nodes), in which all of the connections are weighted.
These networks are used to study nonlinear and complex relations between input and output data.
One of the most well-known unsupervised neural networks in astronomy is a Kohonen self-organizing map (also called self-organizing map, or SOM).
SOMs visualize complex data~\citep{Kohonen82} and show simple geometrical relationships in non-linear high dimensional data~\citep{Kohonen98}.
The result of a SOM is a 1D or 2D network of neurons, which shows the positions of clusters and their relative distance.
Since the 1990s, many studies have utilized SOMs for object classification and clustering (e.g. classifying quasars' spectra, star/galaxy classifications, gamma-ray burst clustering and light curve classification) and photometric redshift estimation~\citep[e.g.][]{Odewahn92, Hernandez94, Murtagh95, Maehoenen95,Scaringi09,Geach12,Fustes13,Meusinger16}.

The K-means algorithm, SOMs, and hierarchical clustering are the main unsupervised methods that are used in astronomical studies~\citep[e.g.][]{DAbrusco12, Aycha16}.
For K-means algorithms, the user must define the number of clusters and for and SOM algorithms the user defines the number of neurons, and the algorithms decide how to separate the data into clusters.
In the hierarchical clustering method, the user must define dissimilarity between the groups; the algorithm combines (or divides) existing groups based on their dissimilarity and creates a hierarchical structure. 
Comparing the SOMs, K-means and hierarchical clustering methods shows that in some cases the hierarchical clustering method mis-classifies the data~\citep[][and references therein]{Mangiameli96}.
We chose the SOM method over the K-means method due to the fact that SOMs not only cluster data, but also show similarities and dissimilarities between the clusters.
Therefore, we can cluster our sample data and study the underlying structure simultaneously.

@@

Using this method, as expected, we are able to divide the mock sample into 4 distinct groups. Their boundaries are shown by grey colours in Fig.~\ref{fig: sample}: regions with high SFR and high dust content, regions with low SFR and high dust content, regions with high SFR and low dust content, and regions with low SFR and dust content. 
In Fig.~\ref{fig: sample}, the upper part belongs to regions with low dust content, while the lower part belongs to regions with high dust content.
The left part of the plot is where regions with low SFR belong and the right side is for high SFR regions.
Grey to black colours show the border between regions.
This network is considered to be a trained network, and can be used to cluster any new data set with similar entries.

@@

     \cite{Dim15} used {\it Spitzer}/IRS observations to study PAHs in 10 regions chosen to span a range of mid-infrared flux ratios in M31 (Fig.~\ref{fig: regions in m31}). 
     Fluxes and equivalent widths of PAH and atomic line features were measured using the~\textsc{pahfit idl} tool~\citep{Smith07b}.
     Metallicity for all regions except regions 5 and 8 was adopted from~\citep{Sanders12} and for regions 5 and 8, \cite{Dim15} estimated metallicity from the radial metallicity profile.
     \cite{Dim15} also calculated the radiation hardness index (RHI) for these 10 regions using $\left (\log \frac{[Ne~\textsc{iii}]}{[Ne~\textsc{ii}]} + \left [ 0.71 + 1.58\log\frac{[S~\textsc{iV}]}{[S~\textsc{iii}]} \right ] \right )/2$ \citep{}. %PB: NEED CITATION HERE
     We used the measured PAH feature fluxes, metallicity and RHI of each region as the input for self-organizing maps.

>>>>

Some of the most tested methods of data mining in astronomy are artificial neural networks (ANNs)~\citep[e.g.][and references therein]{Hossein14,Hossein16}.
ANNs are inspired to work in the same way that the neurons work in a human brain.
These are networks of interconnected neurons (nodes), in which all of the connections are weighted.
These networks are used to study nonlinear and complex relations between input and output data.
There are quite a few data mining methods available, and each one of them has its own advantages or disadvantages. 
One of the most well-known unsupervised neural networks in astronomy is a Kohonen self-organizing map (also called self-organizing map, or SOM).
SOMs visualize complex data~\citep{Kohonen82} and show simple geometrical relationships in non-linear high dimensional data~\citep{Kohonen98}.
The result of a SOM is a 1D or 2D network of neurons, which shows the positions of clusters and their relative distance.
The main disadvantages of this method are not handling errors in the data points and having relative results.
However, the one of the main advantages of this method beside the mentioned ones is that this method is well-tested in Astronomy. 
Since the 1990s, many studies have utilized SOMs for object classification and clustering (e.g. classifying quasars' spectra, star/galaxy classifications, gamma-ray burst clustering and light curve classification) and photometric redshift estimation~\citep[e.g.][]{Odewahn92, Hernandez94, Murtagh95, Maehoenen95,Scaringi09,Geach12,Fustes13,Meusinger16}.

The K-means algorithm, SOMs, and hierarchical clustering are the main unsupervised methods that are used in astronomical studies~\citep[e.g.][]{DAbrusco12, Aycha16}.
For K-means algorithms, the user must define the number of clusters and for and SOM algorithms the user defines the number of neurons, and the algorithms decide how to separate the data into clusters.
In the hierarchical clustering method, the user must define dissimilarity between the groups; the algorithm combines (or divides) existing groups based on their dissimilarity and creates a hierarchical structure. 
There are reports of mis-classifies the data~\citep[][and references therein]{Mangiameli96} by the hierarchical clustering method, when it was compared with SOMs and K-means using the same datasets.
We chose the SOM method over the K-means method due to the fact that SOMs not only cluster data, but also show similarities and dissimilarities between the clusters.
Therefore, we can cluster our sample data and study the underlying structure simultaneously.
%The other point is that as long as one of the data mining method works on a dataset, the other method will work on it, too.

@@

Using this method, as expected, we are able to divide the mock sample into 4 distinct groups. Their boundaries are shown by grey colours in Fig.~\ref{fig: sample}: regions with high SFR and high dust content, regions with low SFR and high dust content, regions with high SFR and low dust content, and regions with low SFR and dust content. 
In Fig.~\ref{fig: sample}, the upper part belongs to regions with low dust content, while the lower part belongs to regions with high dust content.
The left part of the plot is where regions with low SFR belong and the right side is for high SFR regions.
Grey to black colours show the border between regions.
As mentioned in Section~\ref{sec: algorithm_somz}, the grey cycle shows the differences between the weight of each neurons and the initial weight is assigned randomly.
Therefore, the colours between neurons shows the relative differences not the absolute ones.
This network is considered to be a trained network, and can be used to cluster any new data set with similar entries.

@@

     \cite{Dim15} used {\it Spitzer}/IRS observations to study PAHs in 10 regions chosen to span a range of mid-infrared flux ratios in M31 (Fig.~\ref{fig: regions in m31}). 
     Fluxes and equivalent widths of PAH and atomic line features were measured using the~\textsc{pahfit idl} tool~\citep{Smith07b}.
     Metallicity for all regions except regions 5 and 8 was adopted from~\citep{Sanders12} and for regions 5 and 8, \cite{Dim15} estimated metallicity from the radial metallicity profile.
     \cite{Dim15} also calculated the radiation hardness index (RHI) for these 10 regions using $\left (\log \frac{[Ne~\textsc{iii}]}{[Ne~\textsc{ii}]} + \left [ 0.71 + 1.58\log\frac{[S~\textsc{iV}]}{[S~\textsc{iii}]} \right ] \right )/2$~\citep{Engelbracht08}. 
     We used the measured PAH feature fluxes, metallicity and RHI of each region as the input for self-organizing maps.
